Introduction

### 2025-01-25 data.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

---
title: "Project Setup"
engine: knitr
---

We will start by setting up a simple example of a reproducible report.

## Create Quarto Project

First, we will need to create a new Quarto project.

If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.

![](images/setup_01.png){width=500px fig-align="center"}

Here, select _New Directory_

![](images/setup_02.png){width=600px fig-align="center"}

And choose the project type _Quarto Project_.

![](images/setup_03.png){width=600px fig-align="center"}

Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.

As we will use Git to track the version history of files, be sure to check _Create a git repository_. If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".

::: {.column-margin}
![renv: A dependency management toolkit for R](images/renv.svg){width=250px}
:::

Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on. Using it makes it easier for others to view and obtain them at a later point in time. Therefore make sure that the box _Use renv with this project_ is checked. Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".

If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.

![](images/setup_04.png){width=600px fig-align="center"}

Click on _Create Project_. Your RStudio window should now look similar to this:

![](images/setup_05.png){.lightbox fig-align="center" alt="The project `code-publishing-exercise` opened in RStudio. The source pane to the top left has a Quarto file open called \"code-publishing-exercise.qmd\". The console pane to the bottom left indicates by its output that renv is active in the current project. The environment pane to the top right indicates that the environment is currently empty. The output pane to the bottom right shows the files in the current project."}

If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.

Make sure that your project is in a consistent state according to `renv` by running:

```{.r filename="Console"}
renv::status()
```

If it reports packages that are not used, synchronize the lock file using:

```{.r filename="Console"}
renv::snapshot()
```

::: {#nte-terminal .callout-note collapse="true"}
### Without RStudio

Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:

```{.bash filename="Terminal"}
quarto create project default code-publishing-exercise
cd code-publishing-exercise/
rm code-publishing-exercise.qmd
git init
git checkout -b main
```

Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Then, initialize `renv`:

```{.r filename="Console"}
renv::init()
```
:::

You are now ready to stage and commit your files. You can either stage files separately or the whole project folder at once. If you do the latter, we recommend you to inspect the untracked changes before staging all of them:

::: {.column-margin}
In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
:::

```{.bash filename="Terminal"}
git status
```

Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file. If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:

```{.bash filename="Terminal"}
git add .
git commit -m "Initial commit"
```

If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file. If you don't check your changes before committing, you might accidentally commit something you'd rather not.

::: {#tip-author-identity-unknown .callout-tip}
If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:

```{.bash filename="Terminal"}
git config user.name "YOUR NAME"
git config user.email "YOUR EMAIL ADDRESS"
```

Then, commit again.
:::

## Decide on Structure

Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it. In general, the directory structure should facilitate understanding a project by breaking it into logical chunks. There is no single best solution, as a good structure depends on where a project's complexity lies. However, it is usually helpful if the different files and folders reflect the execution order. For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders. Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.

Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021]. We recommend you to follow the standards of your field.

For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.

## Add Data

```{r}
#| echo: false

write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
```

You can now download the data set we have prepared for you and put it into your project folder: [`data.csv`](data.csv){download=""}

@@

The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica. It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

@@

Everything you put into the project folder will be shared publicly. For reasons of reproducibility, this should include the data you analyze. Of course, you should only share them to the extent you are allowed to, taking into account:

1. applicable privacy laws (e.g., the GDPR for European citizens),
2. contractual obligations (e.g., with your data provider),
3. copyright of the data and their particular structure, and
4. any _sui generis_ database right.

Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from. For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).

Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are. Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database. As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
:::

@@

Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, or contact information. In this case, as the data have already been published, we only store a reference to its source:

```{r, class.output = "r code-overflow-wrap", echo = FALSE}

dat_source <- citation("palmerpenguins", auto = TRUE) |>
  format(bibtex = FALSE, style = "text")
dat_source |>
  gsub("\\n", " ", x = _) |>
  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
  paste0("dat_source <- \"", ... = _, "\"") |>
  cat()
```

@@

For a more elaborate introduction to `pointblank`, you can read their [Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html).

@@

```{r, class.output = "r code-overflow-wrap", echo = FALSE}

@@

Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released. As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information. We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/). Finally, the package is used to create a metadata file that contains the schema.

@@

Having added the data and its documentation, one can view and record the utilized packages with `renv`...

@@

...and go through the commit routine:

## Add Code

In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file. The manuscript contains code together with a written narrative. Download the two files to your computer and put them into your project folder.

- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}

The manuscript explores differences in bill length between male and female penguins, feel free to read through it.

::: {#imp-take-copyright-seriously .callout-important}
### Take Copyright Seriously

If you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so. Only build on existing work for which you receive an express grant of relevant rights. How do you know you are allowed to copy, edit, and share the two files linked above?
:::

:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
### Hint for "Take Copyright Seriously"

Have a look at the [about page](about.qmd).
::::

As the manuscript uses some new packages, install them with:

```{.r filename="Console"}
renv::install()
```

The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020]. It can be installed in the project using the following command:

```{.bash filename="Terminal"}
quarto add --no-prompt wjschne/apaquarto
```

::: {#tip-apaquarto .callout-tip}
#### Not a Psychologist?

If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.

Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:

```{.yml filename="Manuscript.qmd"}
format:
  pdf:
    pdf-engine: lualatex
    documentclass: scrartcl
    papersize: a4
```
:::

Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents. A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:

```{.bash filename="Terminal"}
quarto install tinytex
```

You should now be able to render the document using Quarto:

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```

This should create a PDF file called `Manuscript.pdf` in your project folder.

::: {#tip-update-quarto .callout-tip}
If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
:::

With the code being added, one can use `renv` again to view and record the new packages:

::: {#tip-renv-status .callout-tip}
Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
:::

Finally, make your changes known to Git:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add manuscript"
```

::: {#wrn-credentials .callout-warning}
### Beware of Credentials

Sometimes, a data analysis requires the interaction with online services:

- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
- an analysis may be conducted with the help of AI providers.

In these cases, make sure that the code you check in to Git does not contain any credentials that are required for accessing these services. Instead, make use of environment variables which are defined in a location that is excluded from version control. When programming with R, you can define them in a file called `.Renviron` in the root of your project folder:

```{.ini filename=".Renviron"}
MY_FIRST_KEY="your_api_key_here"
MY_SECOND_KEY="your_api_key_here"
```

When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:

```r
query_api(..., api_key = Sys.getenv("MY_FIRST_KEY"))
```

Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that contains credentials, you can follow @Chacon2024.
:::

### Coding Best Practices

Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.

- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root. For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.

- __Keep it simple.__ Add complexity only when you must. Whenever there's a boring way to do something and a clever way, go for the boring way. If the code grows increasingly complex, refactor it into separate functions and files.

- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.

- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.

- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).

- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future. Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.]. We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.

- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.

- __Test your code.__ Test your code with scenarios where you know what the result should be. Turn bugs you discovered into test cases. Use linting tools^[A linting tool analyzes your code without actually running it. Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).

- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality. R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/). Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013. And even if you don't follow a style guide, be consistent.

This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:

- "Tidy design principles" [@Wickham2023Design]
- "The Good Research Code Handbook" [@Mineault2021]
- "Quality assurance of code for analysis and research" [@UKGovernment2020]
- "The Art of UNIX Programming" [@Raymond2003]

> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
> 
> --- @Fowler1999, p. 15

## Cite Data and Software

If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.

### Data

Put simply, all data relied upon should be cited to allow for precise identification and access. From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):

@@

Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?

::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
#### Hint for citing the data set

As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:

```{r, echo = -1}
invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
citation("palmerpenguins")
```

As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).

Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`. Then, add a sentence to the manuscript such as follows:

```md
The analyzed data are by @horst2020.
```

@@

### Software

When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies. You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not. As with data, citations should allow for exact identification and access. From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):

> **1\. Importance**: Software should be considered a legitimate and citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be included in the metadata of the citing work, for example in the reference list of a journal article, and should not be omitted or separated. Software should be cited on the same basis as any other research product such as a paper or a book, that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
> 
> **5\. Accessibility**: Software citations should facilitate access to the software itself and to its associated metadata, documentation, data, and other materials necessary for both humans and machines to make informed use of the referenced software.
> 
> **6\. Specificity**: Software citations should facilitate identification of, and access to, the specific version of software that was used. Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.

In practice, the first step is to identify all pieces of software the project relies on. A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before. Then there are the individual R packages, Quarto extensions, and $\TeX$ packages. All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper. For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit. A system dependency is additional software that you require on your computer apart from the R package.

::: {#fig-software-citation}
```{mermaid}
flowchart TB
  asks_for_citation("Does the software<br>ask you to cite it?")
  critical_or_unique_contribution("Did the software<br>play a critical part<br>in or contributed<br>something unique<br>to your research?")
  manipulate("Did the software<br>manipulate or create<br>your data, software,<br>or other outputs?")
  relies_on_credit("Do the authors of<br>the software rely<br>on academic credit<br>for funding?")
  cite[Cite!]
  nocite[Don't cite!]

  asks_for_citation --"Yes"--> cite
  asks_for_citation --"No"--> critical_or_unique_contribution
  critical_or_unique_contribution --"Yes"--> cite
  critical_or_unique_contribution --"No"--> manipulate
  manipulate --"Yes"--> cite
  manipulate --"No"--> relies_on_credit
  relies_on_credit --"Yes"--> cite
  relies_on_credit --"No"--> nocite
```

"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Simplified from original.
:::

Now, add references for the software you would like to cite to the manuscript. In the following, we will demonstrate this for R and all R packages by using the R package `grateful`. For arbitrary software, you can use the [CiteAs service](https://citeas.org/) to create appropriate citations.

Add the following code chunk to the end of the discussion in the manuscript:

``````{.qmd filename="Manuscript.qmd"}
```{{r}}
#| echo: false

grateful::cite_packages(
    output = "paragraph",
    out.dir = ".",
    omit = NULL,
    dependencies = TRUE,
    passive.voice = TRUE,
    bib.file = "grateful-refs"
)
```
``````

This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that this automatic detection can miss packages in some circumstances, therefore always verify the rendered result.] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:

```{.yml filename="Manuscript.qmd"}
bibliography:
  - Bibliography.bib
  - grateful-refs.bib
```

Use `renv` to view, install, and record the newly used package:

```{.r filename="Console"}
renv::status()
renv::install()
renv::snapshot()
```

Finally, render the document again and commit the changes:

```{.bash filename="Terminal"}
quarto render Manuscript.qmd

git status
git add .
git commit -m "Cite data and software"
```

## The Last Mile

`renv` only records the versions of R packages and of R itself. This means that everything we have not decided to cite in the previous step is not documented anywhere. We will cover system dependencies when [creating a README](make_readme.qmd). For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies). Do run the following:

```{.bash filename="Terminal"}
quarto use binder
```

This will create a few additional files which facilitate reconstructing the computational environment in the future.^[Either using [repo2docker](https://repo2docker.readthedocs.io/) or the public [binder](https://mybinder.org/) service.] As always, commit your changes:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add repo2docker config"
```

You are now all set up to prepare your project for sharing!

>>>>

---
title: "Add Data"
engine: knitr
---

## Add Data File

```{r}
#| include: false

invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
stopifnot(packageDescription("palmerpenguins", fields = "License") == "CC0")
write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
```

You can now download the data set we have prepared for you and put it into your project folder:

[data.csv](data.csv){.btn .btn-lg .btn-info download=""}

@@

The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica. It was made available by Allison Horst, Alison Hill, and Kristen Gorman under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

@@

Everything you put into the project folder will be shared publicly. For reasons of reproducibility, this should include the data you analyze. Of course, you should only share them to the extent you are allowed to. Besides copyright and similar rights, you need to take into account applicable privacy laws (e.g., the GDPR for European citizens) and contractual obligations (e.g., with your data provider).

Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from. For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu). Another resource to look at is the presentation "Creating reproducible packages when data are confidential" by @Vilhuber2024.
:::

## Add Data Dictionary

@@

Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, applied preprocessing steps, or contact information. In our case, as the data have already been published, we only store a reference to its source.

As the data set is from the R package `palmerpenguins`, if you have it installed you can use the function `citation()` to create such a reference:

```{r, filename="create_data_dictionary.R"}
dat_source <- citation("palmerpenguins", auto = TRUE) |>
  format(bibtex = FALSE, style = "text")
```

Without the package `palmerpenguins` installed, you can find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation) and run the following:

```{r}
#| echo: false
#| class-output: "r code-overflow-wrap"
#| attr-output: 'filename="create_data_dictionary.R"'

# We have provided the data set as CSV file to the readers.
# Therefore, we cannot assume or require that readers have
# the R package palmerpenguins installed. Instead, we create
# the citation on our end and hide how we obtained it.
dat_source |>
  gsub("\\n", " ", x = _) |>
  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
  paste0("dat_source <- \"", ... = _, "\"") |>
  cat()
```

@@

For a more elaborate introduction to `pointblank`, you can read its "[Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html)".

@@

```{r}
#| echo: false
#| class-output: "r code-overflow-wrap"
#| attr-output: 'filename="Console"'

@@

Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released. As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information. We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under CC0\ 1.0. Finally, the package is used to create a metadata file that contains the schema.

@@

Having added the data and its documentation, one can view and record the utilized packages with `renv`:

@@

## Add Data Citation and Attribution

All data relied upon should be cited in the manuscript to allow for precise identification and access. From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):

@@

Now, it's your turn to add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?

::: {#tip-cite-palmerpenguins .callout-tip collapse="true"}
#### Citing the Data Set (Solution)

You can find an appropriate BibTeX entry on [the package website](https://allisonhorst.github.io/palmerpenguins/#citation) or with the function `citation()`:^[Note that this function requires to have the respective package installed.]

```{r}
#| class.output: "bib code-overflow-scroll"
#| attr-output: 'filename="Bibliography.bib"'

citation("palmerpenguins", auto = TRUE) |>
  transform(key = "horst2020") |>
  toBibtex()
```

Copy the BibTeX entry to the file `Bibliography.bib`. Then, find the line in the manuscript that says "cite data here" and replace it with a sentence such as the following:

```{.md filename="Manuscript.qmd"}
The analyzed data are by @horst2020.
```

@@

While citation happens in the manuscript for reasons of academic integrity and reproducibility, to comply with any licenses you also may need to provide attribution within your project folder. Even though the data file we use here does not require attribution, we recommend adding a short paragraph to `LICENSE.txt`:

```{r}
#| echo: false
#| class-output: "txt code-overflow-wrap"
#| attr-output: 'filename="LICENSE.txt"'

cat(paste0(
  "The penguins data stored in \"data.csv\" by Allison Horst, Alison Hill, and Kristen Gorman available from <",
  citation("palmerpenguins", auto = TRUE)$url,
  "> are licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>"
))
```

As before, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).

Finally, you can go through the commit routine:

### 2025-01-25 intro.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

In the following, we will provide a brief introduction to the concept of _research compendia_.

@@

@@

This tutorial not only covers sharing the source code, but also connecting it to the results through the creation of dynamic documents. Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article. Dynamic documents bundled together with any necessary data and auxiliary software are called a [research compendium](https://research-compendium.science/) [@Gentleman2007].

@@

When creating a research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:

>>>>

In the following, we will set the stage by highlighting the importance of sharing all materials, documenting their usage, and linking results with their computations, leading to the creation _research compendia_.

@@

## Document Materials

However, there's more to code publishing than sharing. Researchers should document (i.e., track) which data, software, images, texts, and other materials contributed to their work. In more elaborate terms, researchers should preserve the _provenance_ of shared materials, for at least three tangible reasons:

- **Academic integrity**: Providing citations for works that are not one's own avoids plagiarism.
- **Complying with the law**: Providing attribution and license texts (among other things) may be a legal obligation when redistributing materials.
- **Facilitating reproductions**: Stating utilized software and data with their exact versions helps reproducers.

Of course, when and how to reserve provenance varies for these reasons. Rephrasing someone's text avoids issues of copyright, but may be plagiarism. And bibliographies of various styles may comply with scientific citation standards, but if they omit the first names of authors or the version number of computer programs, they might not help with matters of copyright or reproducibility.

In this tutorial, we'll consider all of these reasons important. And there are even more: Documenting the provenance puts research into context and allows others to understand how it came about. Also, cited authors benefit from the citation as they gain potential readers.

@@

For the purpose of this tutorial, tracking the provenance of results deserves particular attention. This means connecting them to the source code through the creation of dynamic documents. Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article. Dynamic documents bundled together with any necessary data and auxiliary software are called [research compendia](https://research-compendium.science/) [@Gentleman2007].

@@

To recap, in this tutorial you will share your materials, document their usage, and connect results to the underlying source code through the creation of a research compendium. Make sure to consider the following things along the way [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
